Variational Monte Carlo on bosonic systems

### 2018-02-27 project_1/article/article.tex
Replace tab with spaces
Remove explicit expressions
Add ratio test to algorithm

    \begin{widetext}
        For the Laplacian we switch a little back and forth between the
        two ways of representing the $\nabla_k$-operator. We thus get
        \begin{align}
            \nabla_k^2 u_{km}
            &=
            \frac{\nabla_k \vf{r}_k}{r_{km}}\dpd[]{u_{km}}{r_{km}}
            + \Biggl[
                \nabla_k \frac{1}{r_{km}}
            \Biggr]
            (\vf{r}_k - \vf{r}_m)\dpd[]{u_{km}}{r_{km}}
            + \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
            \nabla_k \dpd[]{u_{km}}{r_{km}} \\
            &= \frac{d}{r_{km}}\dpd[]{u_{km}}{r_{km}}
            - \frac{(\vf{r}_k - \vf{r}_m)^2}{r_{km}^3}
            \dpd[]{u_{km}}{r_{km}}
            + \frac{(\vf{r}_k - \vf{r}_m)^2}{r_{km}^2}
            \dpd[2]{u_{km}}{r_{km}} \\
            &=
            \frac{d - 1}{r_{km}}\dpd[]{u_{km}}{r_{km}}
            + \dpd[2]{u_{km}}{r_{km}},
        \end{align}
        where $d$ is again the dimensionality of the problem. In total
        we can state an intermediate version of the Laplacian occuring
        in the local energy as
        \begin{align}
            \frac{\nabla_k^2\Psi_T(\vf{r})}{\Psi_T(\vf{r})}
            &=
            \frac{\nabla_k^2\phi_k}{\phi_k}
            + 2\frac{\nabla_k\phi_k}{\phi_k}
            \sum_{m \neq k}^N
            \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
            \dpd[]{u_{km}}{r_{km}}
            + \sum_{m\neq k}^N
            \Biggl(
                \frac{d - 1}{r_{km}}\dpd[]{u_{km}}{r_{km}}
                + \dpd[2]{u_{km}}{r_{km}}
            \Biggr)
            \nonumber \\
            &\qquad
            +
            \sum_{m, n \neq k}^N
            \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
            \frac{\vf{r}_k - \vf{r}_n}{r_{kn}}
            \dpd[]{u_{km}}{r_{km}}
            \dpd[]{u_{kn}}{r_{kn}}.
        \end{align}
    \end{widetext}
    Moving on to the derivatives of the interaction terms, $u_{km}$, to
    get an explicit expression for the Laplacian.
    \begin{align}
        \dpd[]{u_{km}}{r_{km}}
        &=
        \frac{a}{r_{km}(r_{km} - a)},
        \\
        \dpd[2]{u_{km}}{r_{km}}
        &= \frac{a^2 - 2ar_{km}}{r_{km}^2(r_{km} - a)^2}.
    \end{align}

\section{Algorithms}

In the project we rely on a Monte Carlo approach of random sampling to obtain numerical results.
Ironic, is it not? To rely randomness to solve problems that must be naturally deterministic?
We simulate random walks over a volume in order to find optimal parameters in our trial wavefunctions.
The most common of such methods, which we make use of herein, is the Metropolis-Hastings algorithm.

\subsection{Metropolis-Hastings Algorithm}
The Metropolis-Hastings algorithm can in our particular situation be condensed down to the following steps,
\begin{enumerate}
    \item The system is initialised by a certain number $N$ of randomly generated positions, or particles. This allows us to evaluate the
     wavefunction at these points and compute the local energy $E_L$.
    \item The initial configuration is changed by setting a new position for one of these particles. The particle is picked at random.
    \item A ratio between new wavefunction density and the previous (initial) density is computed and compared to a random number.
    This acceptance probability decides if the particle move is rejected or accepted.
    The particle is only allowed to move a predetermined step length.
    \item If the particle movement is accepted and the local energy $E_L$ is computed for the new system.
    \item Repeat steps until convergence and an optimum is reached.
\end{enumerate}

The algorithm described above can be applied in an "exhaustive" search of the parameter space in order to find the optimal parameters.
Whether a proposed move is accepted or not is determined by a transition probability and the acceptance probability.
The strength of the algorithm is that the transition algorithm need not be known.

>>>>

    \begin{widetext}
        For the Laplacian we switch a little back and forth between the
        two ways of representing the $\nabla_k$-operator. We thus get
        \begin{align}
            \nabla_k^2 u_{km}
            &=
            \frac{\nabla_k \vf{r}_k}{r_{km}}\dpd[]{u_{km}}{r_{km}}
            + \Biggl[
                \nabla_k \frac{1}{r_{km}}
            \Biggr]
            (\vf{r}_k - \vf{r}_m)\dpd[]{u_{km}}{r_{km}}
            + \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
            \nabla_k \dpd[]{u_{km}}{r_{km}} \\
            &= \frac{d}{r_{km}}\dpd[]{u_{km}}{r_{km}}
            - \frac{(\vf{r}_k - \vf{r}_m)^2}{r_{km}^3}
            \dpd[]{u_{km}}{r_{km}}
            + \frac{(\vf{r}_k - \vf{r}_m)^2}{r_{km}^2}
            \dpd[2]{u_{km}}{r_{km}} \\
            &=
            \frac{d - 1}{r_{km}}\dpd[]{u_{km}}{r_{km}}
            + \dpd[2]{u_{km}}{r_{km}},
        \end{align}
        where $d$ is again the dimensionality of the problem. In total
        we can state an intermediate version of the Laplacian occuring
        in the local energy as
        \begin{align}
            \frac{\nabla_k^2\Psi_T(\vf{r})}{\Psi_T(\vf{r})}
            &=
            \frac{\nabla_k^2\phi_k}{\phi_k}
            + 2\frac{\nabla_k\phi_k}{\phi_k}
            \sum_{m \neq k}^N
            \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
            \dpd[]{u_{km}}{r_{km}}
            + \sum_{m\neq k}^N
            \Biggl(
                \frac{d - 1}{r_{km}}\dpd[]{u_{km}}{r_{km}}
                + \dpd[2]{u_{km}}{r_{km}}
            \Biggr)
            \nonumber \\
            &\qquad
            +
            \sum_{m, n \neq k}^N
            \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
            \frac{\vf{r}_k - \vf{r}_n}{r_{kn}}
            \dpd[]{u_{km}}{r_{km}}
            \dpd[]{u_{kn}}{r_{kn}}.
        \end{align}
    \end{widetext}
    Moving on to the derivatives of the interaction terms, $u_{km}$, to
    get an explicit expression for the Laplacian.
    \begin{align}
        \dpd[]{u_{km}}{r_{km}}
        &=
        \frac{a}{r_{km}(r_{km} - a)},
        \\
        \dpd[2]{u_{km}}{r_{km}}
        &= \frac{a^2 - 2ar_{km}}{r_{km}^2(r_{km} - a)^2}.
    \end{align}
    The local energy and the drift force can now be found by combining these
    expressions. For brevity, we will not write out the explicit expressions as
    these will be called by separated functions in our programs.

\section{Algorithms}
    In the project we rely on a Monte Carlo approach of random sampling to
    obtain numerical results.  We simulate random walks over a volume in order
    to find optimal parameters in our trial wavefunctions.  The most common of
    such methods, which we make use of herein, is the Metropolis-Hastings
    algorithm.

    \subsection{The Metropolis-Hastings Algorithm}
        The Metropolis-Hastings algorithm can in our particular situation be
        condensed down to the following steps:

        \begin{enumerate}
            \item The system is initialised by a certain number $N$ of randomly
                generated positions, or particles. This allows us to evaluate
                the wavefunction at these points and compute the local energy
                $E_L$.

            \item The initial configuration is changed by setting a new position
                for one of these particles. The particle is picked at random.

            \item A ratio between new wavefunction density and the previous
                (initial) density is computed and compared to a random number.
                This acceptance probability decides if the particle move is
                rejected or accepted.  The particle is only allowed to move a
                predetermined step length.

            \item If the particle movement is accepted and the local energy
                $E_L$ is computed for the new system.

            \item Repeat steps until convergence and an optimum is reached.

        \end{enumerate}

        The algorithm described above can be applied in an "exhaustive" search
        of the parameter space in order to find the optimal parameters.  Whether
        a proposed move is accepted or not is determined by a transition
        probability and the acceptance probability.  The strength of the
        algorithm is that the transition algorithm need not be known. For
        example, the simplest case is to accept the new state, i.e., the new
        position for the random walker, if the ratio
        \begin{align}
            w &= \frac{\Psi_T(\vf{r}_{i + 1})}{\Psi_T(\vf{r}_{i})},
        \end{align}
        where $\vf{r}_{i + 1}$ are all the positions at step $i + 1$, is greater
        than a uniform probability $p \in [0, 1)$.
